### 2020-04-18 covid19env/data_processing/01-Data-Preprocessing.Rmd

---
title: "COVID-19-Panel-Spain"
author: "TURURU"
date: "4/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, collapse=TRUE, message=FALSE}
#library(meteoland)
#library(geosphere)
library(ggthemes)
library(lubridate)
library(readr)
library(sf)
library(tidyverse)
library(units)
#library(aemet)
#library(spdep)
#library(spsur)
#library(spatialreg)
#library(systemfit)
#library(plm)
#library(splm)
```

# Los datos Spain

Algunas notas sobre los datos:

* **Los datos son provisionales**

* Las cifras oficiales de Diagnosticados/Muertes estan disponibles a nivel de Comunidad Autónoma (CCAA). 
Aunque sean oficiales hay dudas sobre ellas.
Los informes están disponibles aquí: 
https://www.mscbs.gob.es/en/profesionales/saludPublica/ccayes/alertasActual/nCov-China/situacionActual.htm

* Algunas regiones (CCAA) ofrencen datos a nivel de provincia, otras no.

* Los datos a nivel provincial, principalmente por el esfuerzo de la televisón española (RTVE).
https://www.rtve.es/noticias/20200407/mapa-del-coronavirus-espana/2004681.shtml

* Estos datos junto con otros recogidos de diversas fuentes están recogidos en GitHub en la web de datalista: https://github.com/datadista/datasets/tree/master/COVID%2019

* Los datos climatológicos (temperaturas, etc) son oficiales 
http://www.aemet.es/

* Los datos de poblacion tb son oficiales del Instituto Nacional de Estadística
http://www.ine.es/

## Lectura datos COVID-19 Spain

Principales características:

* **N=50 provincias**
* **T=22 días, desde el 14 de marzo hasta el 4 de abril**
* **Desde el día 14 Marzo hasta el 4 de Abril** EL CONFINAMIENTO SE INICIÓ EL 16 DE MARZO
* **Empezamos el 14 de Marzo porque es el primer dia con al menos 1 caso en todas las provincias**

```{r}
## Lectura de Datos COVID-19
BD <- read.table(file="COVID-19-Diagnosticados-Extendido.csv",sep = ";",header = TRUE)
BD$X <- NULL
BD$X.1 <- NULL
```

Pivot table:
```{r}
BD <- BD %>%
  rename(province = `ï..Provincia`)

covid19_spain <- BD %>%
  pivot_longer(starts_with("D"),
               names_to = "Date",
               values_to = "Cases") %>%
  mutate(Date = rep(seq(dmy("13-03-2020"), 
                        dmy("13-04-2020"), 
                        by = "days"), 
                    50))
```

## Read GDP per capita

Read data
```{r}
## Read file
GDPpc <- read_csv("PIBpc_Spain.csv", col_names = TRUE) %>%
  select(ID_INE, gdppc16)
```

## Read population and demographic data

```{r}
population.sf <- st_read("Shp/Provincias.shp") %>%
  st_drop_geometry()
```

## Join provincial data to table with COVID-19 cases

Join the tables and convert to simple features:
```{r}
covid19_spain <- covid19_spain %>%
  left_join(population.sf %>% 
              select(ID_INE, POPULATION), 
            by = "ID_INE")
```

Calculate the incidence of COVID-19 cases (cases 100,000 hab):
```{r}
covid19_spain <- covid19_spain %>%
  mutate(Incidence = Cases * 100000/POPULATION) %>%
  select(-POPULATION)
```

## Read sf with provinces

```{r}
provinces_spain <- sf::st_read("Shp/shapefiles_provincias_espana.shp") #
```

Join provincial-level data to sf:
```{r}
provinces_spain <- provinces_spain %>%
  transmute(ID_INE = cartodb_id, geometry) %>%
  left_join(covid19_spain %>% 
              filter(Date == "2020-03-14") %>%
              select(province, 
                     CCAA, 
                     ID_INE),
            by = "ID_INE") %>%
  left_join(population.sf,
            by = "ID_INE") %>%
  left_join(GDPpc, 
            by = "ID_INE")
```

## Select and rename variables

Rename dataframe:
```{r}
provinces_spain <- provinces_spain  %>%
  mutate(Density = POPULATION/units::set_units(st_area(provinces_spain), km^2)) %>% 
  select(province, 
         CCAA,
         ID_INE,
         Population = POPULATION,
         Density,
         Older = MAY_65, 
         Median_Age = EDAD_MED, 
         Male2Female = HOM_MUJ,
         GDPpc = gdppc16,
         Transit = METRO,
         Altitude = ALTITUD,
         Coast = COSTA,
         Meteo_Station = ESTACION_M,
         Area = AREA,
         geometry)
```

## Lectura datos climáticos

* Se ha usado el paquete **meteoland** para obtener las temperaturas de distintas estaciones meteorológicas (el data frame stations lista las estaciones). Para cada provincia de ha elegido una estación meteorológica que fuera representativa y que contuviera datos de temperaturas (Max/Min/Media), Pluviosidad, y horas de sol.

En general estas estaciones no recogen información sobre la humedad relativa (HR). Para poder incluir en el modelo la HR se han considerado dos opciones. La primera alternativa es la estimación en base a las temperaturas máximas y mínimas (ver nota en 'COVID-19-Datos-Climáticos-Spain.R'). la segunda alternativa es recoger datos sobre la altitud de cada provincia (datos de la capital de provincia en wikipedia). La altitud es una proxy de la humedad ya que a mayor altitud el clima es más seco y el grado de humedad menor.

## Process climate data

Load climate data (ignore the humidity data in this file; it was a calculated variable):
```{r}
load("COVID-19-Datos_Climaticos_Spain-Extendido.RData")
```

Read humidity data
```{r}
## Read file
Hum <- read_csv("COVID-19-Humedad-Spain.csv", col_names = TRUE) %>%
  rename(provincia = Provincia) %>% 
  select(provincia:D11Abr)
```

Prepare climatic factors:
```{r}
# Maximum temperature
TMax <- TMax %>% select(-ID, -provincia) %>%
  pivot_longer(starts_with("T"), 
               names_to = "Date",
               values_to = "Max_Temp") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"), 
                        by = "days"), 
                    50))

# Minimum temperature
TMin <- TMin %>% select(-ID, -provincia) %>%
  pivot_longer(starts_with("T"), 
               names_to = "Date",
               values_to = "Min_Temp") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"), 
                        by = "days"), 
                    50))

# Mean temperature
TMed <- TMed %>% select(-ID, -provincia) %>%
  pivot_longer(starts_with("T"), 
               names_to = "Date",
               values_to = "Mean_Temp") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"), 
                        by = "days"), 
                    50))

# Hours of sunshine 
Sun <- Sun %>% select(-ID, -provincia) %>%
  pivot_longer(starts_with("S"), 
               names_to = "Date",
               values_to = "Sunshine_Hours") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"), 
                        by = "days"), 
                    50))

# Precipitation 
Plu <- Plu %>% select(-ID, -provincia) %>%
  pivot_longer(c(starts_with("P")), 
               names_to = "Date",
               values_to = "Precipitation") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"),
                        by = "days"), 
                    50)) %>%
  replace_na(list(Precipitation = 0)) # Replace two missing values in Malaga on 2020-03-01 and 2020-03-02

# Humidity 
Hum <- Hum %>% select(-provincia) %>%
  pivot_longer(c(starts_with("D")), 
               names_to = "Date",
               values_to = "Humidity") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"),
                        by = "days"), 
                    50))
```

Join all climatic factors:
```{r}
climatic_vars <- TMax %>%
  left_join(TMin, by = c("ID_INE", "Date")) %>%
  left_join(TMed, by = c("ID_INE", "Date")) %>%
  left_join(Sun, by = c("ID_INE", "Date")) %>%
  left_join(Plu, by = c("ID_INE", "Date"))  %>%
  left_join(Hum, by = c("ID_INE", "Date"))
```

It is important to note that someone who is diagnosed on day "date" was likely infected between 2 and 12 days earlier. Therefore, the climatic variables need to be with respect to the likely time of contagion. For this reason, we calculate a lagged average using the variable from date-minus-12 to date-minus-5 days. These variables are called "lag8".
```{r}
# Lagged mean temperature: 8 day average
TMed_lag8 <- TMed %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Mean_Temp, 12) + 
                      lag(.x$Mean_Temp, 11)+ 
                      lag(.x$Mean_Temp, 10)+ 
                      lag(.x$Mean_Temp, 9)+ 
                      lag(.x$Mean_Temp, 8)+ 
                      lag(.x$Mean_Temp, 7)+ 
                      lag(.x$Mean_Temp, 6)+ 
                      lag(.x$Mean_Temp, 5))/8) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Mean_Temp_lag8 = value)

# Lagged sunshine hours: 8 day average
Sun_lag8 <- Sun %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Sunshine_Hours, 12) + 
                      lag(.x$Sunshine_Hours, 11)+ 
                      lag(.x$Sunshine_Hours, 10)+ 
                      lag(.x$Sunshine_Hours, 9)+ 
                      lag(.x$Sunshine_Hours, 8)+ 
                      lag(.x$Sunshine_Hours, 7)+ 
                      lag(.x$Sunshine_Hours, 6)+ 
                      lag(.x$Sunshine_Hours, 5))/8) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Sunshine_Hours_lag8 = value)

# Lagged precipitation: 8 day average
Plu_lag8 <- Plu %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Precipitation, 12) + 
                      lag(.x$Precipitation, 11)+ 
                      lag(.x$Precipitation, 10)+ 
                      lag(.x$Precipitation, 9)+ 
                      lag(.x$Precipitation, 8)+ 
                      lag(.x$Precipitation, 7)+ 
                      lag(.x$Precipitation, 6)+ 
                      lag(.x$Precipitation, 5))/8) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Precipitation_lag8 = value)

# Lagged humidity: 8 day average
Hum_lag8 <- Hum %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Humidity, 12) + 
                      lag(.x$Humidity, 11)+ 
                      lag(.x$Humidity, 10)+ 
                      lag(.x$Humidity, 9)+ 
                      lag(.x$Humidity, 8)+ 
                      lag(.x$Humidity, 7)+ 
                      lag(.x$Humidity, 6)+ 
                      lag(.x$Humidity, 5))/8) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Humidity_lag8 = value)
```

A variation of this is a lagged 11-day average, from date-minus-12 to date-minus-2 days. According to [this research](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported) on the incubation period, this is the latency period of 95% of people who develop symptoms. Call this "lag11":
```{r}
# Lagged mean temperature: 8 day average
TMed_lag11 <- TMed %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Mean_Temp, 12) + 
                      lag(.x$Mean_Temp, 11) + 
                      lag(.x$Mean_Temp, 10) + 
                      lag(.x$Mean_Temp, 9) + 
                      lag(.x$Mean_Temp, 8) + 
                      lag(.x$Mean_Temp, 7) + 
                      lag(.x$Mean_Temp, 6) + 
                      lag(.x$Mean_Temp, 5)+ 
                      lag(.x$Mean_Temp, 4)+ 
                      lag(.x$Mean_Temp, 3)+ 
                      lag(.x$Mean_Temp, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Mean_Temp_lag11 = value)

# Lagged sunshine hours: 8 day average
Sun_lag11 <- Sun %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Sunshine_Hours, 12) + 
                      lag(.x$Sunshine_Hours, 11) + 
                      lag(.x$Sunshine_Hours, 10) + 
                      lag(.x$Sunshine_Hours, 9) + 
                      lag(.x$Sunshine_Hours, 8) + 
                      lag(.x$Sunshine_Hours, 7) + 
                      lag(.x$Sunshine_Hours, 6) + 
                      lag(.x$Sunshine_Hours, 5) + 
                      lag(.x$Sunshine_Hours, 4) + 
                      lag(.x$Sunshine_Hours, 3) + 
                      lag(.x$Sunshine_Hours, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Sunshine_Hours_lag11 = value)

# Lagged precipitation: 8 day average
Plu_lag11 <- Plu %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Precipitation, 12) + 
                      lag(.x$Precipitation, 11) + 
                      lag(.x$Precipitation, 10) + 
                      lag(.x$Precipitation, 9) + 
                      lag(.x$Precipitation, 8) + 
                      lag(.x$Precipitation, 7) + 
                      lag(.x$Precipitation, 6) + 
                      lag(.x$Precipitation, 5) + 
                      lag(.x$Precipitation, 4) + 
                      lag(.x$Precipitation, 3) + 
                      lag(.x$Precipitation, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Precipitation_lag11 = value)

# Lagged humidity: 8 day average
Hum_lag11 <- Hum %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Humidity, 12) + 
                      lag(.x$Humidity, 11) + 
                      lag(.x$Humidity, 10) + 
                      lag(.x$Humidity, 9) + 
                      lag(.x$Humidity, 8) + 
                      lag(.x$Humidity, 7) + 
                      lag(.x$Humidity, 6) + 
                      lag(.x$Humidity, 5) + 
                      lag(.x$Humidity, 4) + 
                      lag(.x$Humidity, 3) + 
                      lag(.x$Humidity, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Humidity_lag11 = value)
```

Yet another variation of this is a weighted average 11-day lag, from date-minus-12 to date-minus-2 days, with the weights given by the distribution in Lauer et al. (2020). First, calculate the weights using the parameters reported by Lauer et al.:
```{r}
# Calculate the p-values for the temporal weights
tweights <- data.frame(x = seq(2, 12, 1), p = dlnorm(seq(2, 12, 1), meanlog = 1.621, sdlog = 0.418))

# Sum, to normalize to one
ptotal <- sum(tweights$p)

# Weights are the normalized p-values
tweights <- tweights %>%
  mutate(w = p/ptotal)

tweights <- rbind(data.frame(x = 1, p = 0, w = 0), tweights)
```

Call this "lag11w":
```{r}
# Lagged mean temperature: 8 day average
TMed_lag11w <- TMed %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Mean_Temp, 12) * tweights$w[12] + 
                      lag(.x$Mean_Temp, 11) * tweights$w[11]+ 
                      lag(.x$Mean_Temp, 10) * tweights$w[10]+ 
                      lag(.x$Mean_Temp, 9) * tweights$w[9]+ 
                      lag(.x$Mean_Temp, 8) * tweights$w[8]+ 
                      lag(.x$Mean_Temp, 7) * tweights$w[7]+ 
                      lag(.x$Mean_Temp, 6) * tweights$w[6]+ 
                      lag(.x$Mean_Temp, 5) * tweights$w[5]+ 
                      lag(.x$Mean_Temp, 4) * tweights$w[4]+ 
                      lag(.x$Mean_Temp, 3) * tweights$w[3]+ 
                      lag(.x$Mean_Temp, 2) * tweights$w[2])) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Mean_Temp_lag11w = value)

# Lagged sunshine hours: 8 day average
Sun_lag11w <- Sun %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Sunshine_Hours, 12) * tweights$w[12] + 
                      lag(.x$Sunshine_Hours, 11) * tweights$w[11]+ 
                      lag(.x$Sunshine_Hours, 10) * tweights$w[10]+ 
                      lag(.x$Sunshine_Hours, 9) * tweights$w[9]+ 
                      lag(.x$Sunshine_Hours, 8) * tweights$w[8]+ 
                      lag(.x$Sunshine_Hours, 7) * tweights$w[7]+ 
                      lag(.x$Sunshine_Hours, 6) * tweights$w[6]+ 
                      lag(.x$Sunshine_Hours, 5) * tweights$w[5]+ 
                      lag(.x$Sunshine_Hours, 4) * tweights$w[4]+ 
                      lag(.x$Sunshine_Hours, 3) * tweights$w[3]+ 
                      lag(.x$Sunshine_Hours, 2) * tweights$w[2])) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Sunshine_Hours_lag11w = value)

# Lagged precipitation: 8 day average
Plu_lag11w <- Plu %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Precipitation, 12) * tweights$w[12] + 
                      lag(.x$Precipitation, 11) * tweights$w[11]+ 
                      lag(.x$Precipitation, 10) * tweights$w[10]+ 
                      lag(.x$Precipitation, 9) * tweights$w[9]+ 
                      lag(.x$Precipitation, 8) * tweights$w[8]+ 
                      lag(.x$Precipitation, 7) * tweights$w[7]+ 
                      lag(.x$Precipitation, 6) * tweights$w[6]+ 
                      lag(.x$Precipitation, 5) * tweights$w[5]+ 
                      lag(.x$Precipitation, 4) * tweights$w[4]+ 
                      lag(.x$Precipitation, 3) * tweights$w[3]+ 
                      lag(.x$Precipitation, 2) * tweights$w[2])) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Precipitation_lag11w = value)

# Lagged humidity: 8 day average
Hum_lag11w <- Hum %>% 
  group_by(ID_INE) %>%
  group_modify(~ {((lag(.x$Humidity, 12) * tweights$w[12] + 
                      lag(.x$Humidity, 11) * tweights$w[11]+ 
                      lag(.x$Humidity, 10) * tweights$w[10]+ 
                      lag(.x$Humidity, 9) * tweights$w[9]+ 
                      lag(.x$Humidity, 8) * tweights$w[8]+ 
                      lag(.x$Humidity, 7) * tweights$w[7]+ 
                      lag(.x$Humidity, 6) * tweights$w[6]+ 
                      lag(.x$Humidity, 5) * tweights$w[5]+ 
                      lag(.x$Humidity, 4) * tweights$w[4]+ 
                      lag(.x$Humidity, 3) * tweights$w[3]+ 
                      lag(.x$Humidity, 2) * tweights$w[2])) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(ID_INE, 
            Date = rep(seq(dmy("01-03-2020"),
                           dmy("11-04-2020"), 
                           by = "days"), 
                       50), 
            Humidity_lag11w = value)
```

Join lagged variables to climatic factors:
```{r}
climatic_vars <- climatic_vars %>%
  left_join(TMed_lag8, by = c("ID_INE", "Date")) %>%
  left_join(Sun_lag8, by = c("ID_INE", "Date")) %>%
  left_join(Plu_lag8, by = c("ID_INE", "Date")) %>%
  left_join(Hum_lag8, by = c("ID_INE", "Date")) %>%
  left_join(TMed_lag11, by = c("ID_INE", "Date")) %>%
  left_join(Sun_lag11, by = c("ID_INE", "Date")) %>%
  left_join(Plu_lag11, by = c("ID_INE", "Date")) %>%
  left_join(Hum_lag11, by = c("ID_INE", "Date")) %>%
  left_join(TMed_lag11w, by = c("ID_INE", "Date")) %>%
  left_join(Sun_lag11w, by = c("ID_INE", "Date")) %>%
  left_join(Plu_lag11w, by = c("ID_INE", "Date")) %>%
  left_join(Hum_lag11w, by = c("ID_INE", "Date"))
```

## Join COVID-19 cases and climatic variables

Join the dataframes:
```{r}
covid19_spain <- covid19_spain %>%
  full_join(climatic_vars, by = c("ID_INE", "Date"))
```

Summary of dataframe:
```{r}
summary(covid19_spain)
```

## Plot two control variables Ratio Male/Female; Median_Age

Ratio of male to female in province:
```{r}
ggplot(data = provinces_spain) + 
  geom_sf(aes(fill = Male2Female)) + 
  scale_fill_distiller(palette = "Blues", direction = 1) + 
  theme_tufte()
```

Median age of population in province:
```{r}
ggplot(data = provinces_spain) + 
  geom_sf(aes(fill = Median_Age)) + 
  scale_fill_distiller(palette = "Blues", direction = 1) + 
  theme_tufte()
```

### Incidence progress

Plot the progression of the incidence of COVID-19 cases by province:
```{r}
p1 <- ggplot(data = covid19_spain %>%
               filter(Date == "2020-03-19") %>% 
         left_join(provinces_spain, by = "ID_INE") %>% 
         st_as_sf()) + 
  geom_sf(aes(fill = Incidence)) + 
  scale_fill_distiller(palette = "Reds", direction = 1) + 
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom") + 
  facet_wrap(~ Date, nrow = 3)

p2 <- ggplot(data = covid19_spain %>%
               filter(Date == "2020-03-28") %>% 
         left_join(provinces_spain, by = "ID_INE") %>% 
         st_as_sf()) + 
  geom_sf(aes(fill = Incidence)) + 
  scale_fill_distiller(palette = "Reds", direction = 1) + 
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom") + 
  facet_wrap(~ Date, nrow = 3)

p3 <- ggplot(data = covid19_spain %>%
               filter(Date == "2020-04-03") %>% 
         left_join(provinces_spain, by = "ID_INE") %>% 
         st_as_sf()) + 
  geom_sf(aes(fill = Incidence)) + 
  scale_fill_distiller(palette = "Reds", direction = 1) + 
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom") +
  facet_wrap(~ Date, nrow = 3)

gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

## Save data

Filter missing records:
```{r}
covid19_spain <- covid19_spain %>%
  filter(!is.na(province)) %>% # Remove records that are missing data on the province
  filter(!is.na(Max_Temp)) # Remove records that are missing climatic variables

provinces_spain <- provinces_spain %>%
  drop_na()
```

Summary after cleaning:
```{r}
summary(covid19_spain)
```

Summary after cleaning:
```{r}
summary(provinces_spain)
```

Save:
```{r}
# Dataframe
save(covid19_spain, file = "covid19_spain.RData", compress = "xz")

# Simple features for cartography
save(provinces_spain, file = "provinces_spain.RData", compress = "xz")
```