### 2020-04-18 README.Rmd
Experimenting with equations in README.md

---
output: github_document
always_allow_html: true
bibliography: References.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# A spatial analysis of the environmental correlates of COVID-19 incidence in Spain

Antonio Paez (McMaster University)  
Fernando A. Lopez (Universidad Politecnica de Cartagena)  
Tatiane Menezes (Universidade Federal de Pernambuco)  
Renata Cavalcanti (Universidade Federal de Pernambuco)  
Maira Galdino da Rocha Pitta (Universidade Federal de Pernambuco)

## Abstract

Spreading with astonishing speed, the novel SARS-CoV2 has swept the globe, causing enormous stress to health systems and prompting social distance guidelines and mandates to arrest its progress. While there is encouraging evidence that early public health interventions have slowed the spread of the virus, this has come at a high cost as the global economy is brought to its knees. How and when to ease restrictions to movement hinges in part on the question whether SARS-CoV2 will display seasonality associated with variations in temperature and humidity. In this research, we address this question by means of a spatial analysis of the incidence of COVID-19 in the provinces in Spain. Use of a spatial Seemingly Unrelated Regressions (SUR) approach allows us to model the incidence of reported cases of the disease per 100,000 population, as a function of temperature and humidity, while controlling for GDP per capita, population density, percentage of older adults in the population, and presence of mass transit systems. An interesting aspect of the spatial SUR approach is that it models incidence as a contagion process. Our results indicate that incidence of the disease is lower at higher temperatures. The evidence with respect to humidity is more mixed, with coefficients for this variable that are significant in only some equations. Our control variables also yield interesting insights, as population density and percentage of older adults display negative associations with incidence of COVID-19, whereas the presence of mass transit systems in the province is associated with a greater incidence of the disease.

## Keywords

SARS-CoV2  
COVID-19  
Temperature  
Humidity  
Spatial SUR  
Contagion  
Spain

```{r load-packages, include=FALSE}
# The data package needs to be installed from github repository:
# devtools::install_github("paezha/covid19-environmental-correlates/covid19env")

# This version of spsur may be needed:
# Download spsur_1.0.1.3.tar.gz from https://github.com/paezha/covid19-environmental-correlates and install

library(covid19env)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(lubridate)
#library(meteoland) # Ask Fernando if this package was used in the end
library(plm)
library(sf)
library(spdep)
library(spsur)
library(tidyverse)
library(units)
```

```{r load-data, include=FALSE}
# Load data from package `covid19env`
data("covid19_spain")
data("provinces_spain")
```

```{r data-preparation, include=FALSE}
# Convert GDP per capita to thousands of euros:
provinces_spain <- provinces_spain %>%
  mutate(GDPpc = GDPpc/1000)

# Join provincial data to incidence data and convert to simple features:
covid19_spain <- covid19_spain %>% 
  left_join(provinces_spain %>% st_drop_geometry(),
            by = c("province", "CCAA", "ID_INE"))
```

```{r sort-communities-by-latitude, include=FALSE}
# To visualize the distribution of temperature by CCAA, we sort the communities by latitude, from north to south. We do this by community instead of province because the large number of provinces clutters the plots.

# Dissolve internal provincial boundaries to obtain a simple features object with the autonomous communities: 
ccaa.sf <-  provinces_spain %>% 
  group_by(CCAA) %>% 
  summarize(provinces = n())

# Extract coordinates of autonomous communities
ccaa.coords <- ccaa.sf %>%
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame()

# Join Y coordinate to ccaa.sf
ccaa.sf <- ccaa.sf %>% 
  mutate(long = ccaa.coords$Y)

# Sort autonomous communities from north to south
ccaa.levels <- ccaa.sf %>% 
  arrange(desc(long)) %>% select(CCAA)

ccaa.levels <- as.character(ccaa.levels$CCAA)

# Relevel autonomous communities
covid19_spain <- covid19_spain %>%
  mutate(CCAA = factor(CCAA, levels = ccaa.levels, ordered = TRUE))
```

## Introduction

From a small outbreak linked to a live animal market at the end of 2019 to a global pandemic in a matter of weeks, the SARS-CoV2 virus has threatened to overrun health systems the world over. In efforts to contain the spread, numerous governments in many nations and regions have either recommended or mandated social distancing measures, and as of this writing, millions of people in five continents shelter in place. There are encouraging signs that these measures have arrested the spread of the virus where they have been implemented, and have thus helped to keep a bad situation from becoming even worse [e.g., -@Lancastle2020impact]. However, this has come at a high cost, and the consequences for all spheres of the economy, social, and cultural life have been dire [e.g., @Fernandes2020economic; @Luo2020how]. As a result, there is a sense of urgency to anticipate the progression of the pandemic, in order to plan for progressive lifting of restrictions to movement and social contact [e.g., @Kissler2020projecting]. Needless to say, this has become a delicate, and politically charged, balancing act between public health and the economy [@Gong2020balance].

A salient question in the debate on how and when to ease social distancing measures is whether the virus will display seasonal variations. Earlier, diverse studies have shown the effect of temperature and humidity on the incidence of influenza [e.g., @Makinen2009cold; @Jaakkola2014decline; @Kudo2019low]. Jaakkola et al. [-@Jaakkola2014decline], for example, found that a decrease of temperature and absolute humidity precedes the onset of symptoms of influenza A and B viruses by 3 days in places where the temperature is low. After the 2002-2004 outbreak of SARS, researchers also began to investigate the relationship between these factors and SARS-CoV. In this way, Casanova et al. [-@Casanova2010effects] used two surrogates, namely the gastroenteritis (TGEV) and mouse hepatitis viruses (MHV), to find that virus inactivation was more rapid at temperatures of 20C than 4C, and at 40C than 20C; in terms of humidity, these researchers reported that survival of the virus was lower at moderate relative humidity levels. In a similar vein, but working directly with SARS-CoV, Chan et al. [-@Chan2011effects] found that viability of the virus was lost at temperatures higher than 38C and relative humidity superior to 95%.

While existing research on similar pathogens suggests that SARS-CoV is more stable and potentially easier to transmit in conditions of low temperature and low humidity, it is far from certain that this will also be the case with the novel SARS-CoV2. If such is the case, there is the possibility of easing restrictions to social contact as the weather warms; however, weeks or possibly months of costly measures could become undone if not, and the restrictions are lifted prematurely. Not surprisingly, given the stakes involved, this issue has already triggered a lively debate.

Some of what is thought about the possible seasonality of COVID-19 is based on analogies to the patterns of other known respiratory viruses. However, de Ángel Solá et al. [-@deangel2020weathering] note that "not all seasonal respiratory viruses experience the same spatiotemporal patterns" (section 4). This urges caution when extrapolating from known viruses, and indeed, the evidence in this respect is inconclusive. At a global scale, whereas de Ángel Solá et al. [-@deangel2020weathering] see less risk in the Caribean Basin, Coelho et al. [@Coelho2020exponential] warn that at least during the exponential phase, expansion of the virus is not driven by climate. Similarly, whereas Araujo and Naimi [-@Araujo2020spread] argue that spread of SARS-CoV2 will likely be constrained by climate, Harbert et al. [-@Harbert2020spatial] remain unconvinced that spatial modelling can currently discriminate the distribution of the disease on the basis of climate, at least in the United States. To further complicate matters, much of the relevant work has yet to be peer-reviewed and therefore is open to change (see for example the challenge of @Harbert2020spatial to @Araujo2020spread). In the United States, the National Academy of Sciences, Engineering, and Medicine was engaged by the Office of the Executive for guidance on this matter [see @National2020rapid]. Their conclusion summarizes the situation well (see p. 6): "Some limited data support a potential waning of cases in warmer and more humid seasons, yet none are without major limitations...Additional studies as the SARS-CoV-2 pandemic unfolds could shed more light on the effects of climate on transmission."

With the above considerations in mind, our objective with this paper is to contribute to the knowledge basis regarding the spread of COVID-19 and the influence of environmental factors, particularly temperature and humidity. Hence, here we report results from a spatial analysis of incidence of COVID-19 in fifty provinces in Spain. Spain is one of the countries hardest hit by the virus, and enacted lockdown measures on March 16, 2020, in response to a rapidly growing outbreak of COVID-19. We combine data on reported cases of the disease with metereological information, to create a spatio-temporal dataset covering a period of 30 days. We then join this dataset with provincial-level economic and demographic information to act as controls to our key environmental variables. These data are analyzed using a spatial Seemingly Unrelated Regressions (SUR) approach, which allows us to model incidence of COVID-19 as a spatial contagion process.

The results provide evidence of the effect of temperature on the incidence of COVID-19. The evidence concerning humidity is more mixed: while the direction of the effect is negative, as anticipated, the coefficients for this variable are only significant in some of equations in the system. Our control variables also provide some intriguing insights. The results of this analysis provide support to the hypothesis of seasonality of the novel SARS-CoV2, and should be of interest to public health officials and policy makers wrestling with the question of the trajectory of the pandemic.

Please note that this paper is prepared as a reproducible research document. The source R Markdown document, as well as all data and code needed to reproduce/review/extend the analysis can be obtained from the following repository:

https://github.com/paezha/covid19-environmental-correlates

## Context, Data, and Methods

### Covid-19 in Spain

When was the first case reported? When did the country close international arrivals? When did initial measures of social distancing were implemented? The lockdown?

### Selection of Variables

Explain the rationale for selecting the variables.

For example: The literature about COVID-19 suggested that population density is the one of the most important proliferate cause of these viscous, however this ill spread with different intensity at big cities of the world. Controlling for socioeconomic characteristics the objective of this paper is observe the effect of clime on COVID-19  proliferation.

Some questions are important to point out. Is not evident present a diary association between the temprature (resp. humidity) with declarate case. There is evidences (ref) that a time lag is necessary between the day that the individuo inffect and the case is diagnosticate

### Sources of Data

Our dataset includes information about the daily number of cases of COVID-19 reported in Spain \footnote{cases tested} at the provincial level (NUTIII in eurostat terminology) from March 13th to April 11th (30 days and 50 spatial units). The Spanish National Government publishes periodic updates at the regional level (NUTII) and the information is also released at the provincial level as part of a collaborative project. This information is compiled from several sources, mainly the official web pages of the Spanish regional goverments. In addition, we consider two sets of explanatory variables. The first one, and the focus of this research, is set of two environmental variables, namely temperature and humidity. The second set provides some relevant controls for multivariate analysis, and refers to economic and demographic attributes of the province. Table \ref{tab:descriptive-statistics} shows the descriptive statistics and the provencance of the data.

```{r descriptive-statistics, echo=FALSE}
data.frame(Variable = c("COVID-19 Incidence",
                        "GDPpc", 
                        "Older", 
                        "Population Density",
                        "Mean Temperature",
                        "Humidity"),
           Note = c("Incidence in reported cases of SARS-19 per 100,000 people",
                    "GDP per capita in €1,000s",
                    "Percentage of people aged 65 and older in the province",
                    "Population density in the province in people per sq.km",
                    "Mean temperature in province by date, in C",
                    "Relative humidity in province by date"),
           Min = c(min(covid19_spain$Incidence),
                   min(provinces_spain$GDPpc),
                   min(provinces_spain$Older),
                   min(provinces_spain$Density),
                   min(covid19_spain$Mean_Temp),
                   min(covid19_spain$Humidity)),
           Mean = c(mean(covid19_spain$Incidence),
                   mean(provinces_spain$GDPpc),
                   mean(provinces_spain$Older),
                   mean(provinces_spain$Density),
                   mean(covid19_spain$Mean_Temp),
                   mean(covid19_spain$Humidity)),
           Max = c(max(covid19_spain$Incidence),
                   max(provinces_spain$GDPpc),
                   max(provinces_spain$Older),
                   max(provinces_spain$Density),
                   max(covid19_spain$Mean_Temp),
                   max(covid19_spain$Humidity)),
           SD = c(sd(covid19_spain$Incidence),
                   sd(provinces_spain$GDPpc),
                   sd(provinces_spain$Older),
                   sd(provinces_spain$Density),
                   sd(covid19_spain$Mean_Temp),
                   sd(covid19_spain$Humidity)),
           Source = c("Montera34",
                      "INE",
                      "INE",
                      "INE",
                      "AEMET",
                      "MAPA")) %>%
  kable("html",
        booktabs = TRUE,
        digits = 2,
        caption = "\\label{tab:descriptive-statistics}Descriptive statistics") %>%
  kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  footnote(general = c("Montera34: https://code.montera34.com:4443/numeroteca/covid19/-/blob/master/README_providencialdata19.md", 
                       "INE (Instituto Nacional de Estadistica): https://www.ine.es/",
                       "AEMET (Agencia Estatal de Meteorologia): http://eportal.mapa.gob.es",
                       "MAPA (Ministerio de Agricultura, Pesca y Alimentacion): http://eportal.mapa.gob.es"))
```

### Methods: Spatial SUR

Seemingly Unrelated Regressions are used in the context of...

The classical SUR model without spatial effects (from here, SUR-SIM) is a stack of equations as follows:

<a href="https://www.codecogs.com/eqnedit.php?latex=\begin{equation}&space;\label{eq:sur-sim}&space;\begin{bmatrix}&space;y_1&space;\\&space;y_2&space;\\&space;\vdots&space;\\&space;y_T&space;\end{bmatrix}&space;=&space;\begin{bmatrix}&space;X_1&space;&&space;0&space;&&space;\cdots&space;&&space;0&space;\\&space;0&space;&&space;X_2&space;&&space;\cdots&space;&&space;0&space;\\&space;\vdots&space;&&space;\vdots&space;&&space;\ddots&space;&&space;\vdots&space;\\&space;0&space;&&space;0&space;&&space;\cdots&space;&&space;X_T&space;\end{bmatrix}&space;\&space;\begin{bmatrix}&space;\beta_1&space;\\&space;\beta_1&space;\\&space;\vdots&space;\\&space;\beta_T&space;\end{bmatrix}&space;&plus;&space;\begin{bmatrix}&space;\epsilon_1&space;\\&space;\epsilon_2&space;\\&space;\vdots&space;\\&space;\epsilon_T&space;\end{bmatrix}&space;\end{equation}" target="_blank"><img src="https://latex.codecogs.com/gif.latex?\begin{equation}&space;\label{eq:sur-sim}&space;\begin{bmatrix}&space;y_1&space;\\&space;y_2&space;\\&space;\vdots&space;\\&space;y_T&space;\end{bmatrix}&space;=&space;\begin{bmatrix}&space;X_1&space;&&space;0&space;&&space;\cdots&space;&&space;0&space;\\&space;0&space;&&space;X_2&space;&&space;\cdots&space;&&space;0&space;\\&space;\vdots&space;&&space;\vdots&space;&&space;\ddots&space;&&space;\vdots&space;\\&space;0&space;&&space;0&space;&&space;\cdots&space;&&space;X_T&space;\end{bmatrix}&space;\&space;\begin{bmatrix}&space;\beta_1&space;\\&space;\beta_1&space;\\&space;\vdots&space;\\&space;\beta_T&space;\end{bmatrix}&space;&plus;&space;\begin{bmatrix}&space;\epsilon_1&space;\\&space;\epsilon_2&space;\\&space;\vdots&space;\\&space;\epsilon_T&space;\end{bmatrix}&space;\end{equation}" title="\begin{equation} \label{eq:sur-sim} \begin{bmatrix} y_1 \\ y_2 \\ \vdots \\ y_T \end{bmatrix} = \begin{bmatrix} X_1 & 0 & \cdots & 0 \\ 0 & X_2 & \cdots & 0 \\ \vdots & \vdots & \ddots & \vdots \\ 0 & 0 & \cdots & X_T \end{bmatrix} \ \begin{bmatrix} \beta_1 \\ \beta_1 \\ \vdots \\ \beta_T \end{bmatrix} + \begin{bmatrix} \epsilon_1 \\ \epsilon_2 \\ \vdots \\ \epsilon_T \end{bmatrix} \end{equation}" /></a>

\noindent where $y_{t}=(y_{1t},...,y_{Nt})$ is a $N \times 1$ vector, and in our case $y_{st}$ is the incidence ratio in the province $s$ ($s=1,...,N$) the day $t$ $(t=1,...,T)$; $X_t=(X^1,...,X^{k_t})$ is a $N \times k_t$ matrix of the $k_t$ independent variables, $X_i=(X^i_{st})$; $\beta_t=(\beta_{1t},...,\beta_{Nt})$ is a vector of coefficients and $\epsilon_t=(\epsilon_{1t},...,\epsilon_{Nt})$ is the vector of residuals.

The more relevant characteristic of a SUR model is the dependence structure among the vectors of residuals, namely:

\begin{equation}
\label{eq:sur-err}
E[\epsilon_t \epsilon'_{t'}]=\sigma_{tt'}
\end{equation}

Note that this specification is very flexible, in that it allows changes in the coefficients $\beta_{it}$ in order to modulate the effect of $X^i_{.t}$ on $y_t$. This flexibility can be reduced and it is posible to impose restrinctions considering some $\beta$ coefficients as being constant over time. In this case, we can reformulate the coefficients expression of $\beta_t = (\beta_{1}, \cdots, \beta_{r-1}, \beta_{r}, \beta_{r+1}, \cdots, \beta_{Nt})$ considering the first $r$ coefficients contanst.

Equation \ref{eq:sur-sim}) can be rewriten in compact form:

\begin{equation}
y = X \beta + \epsilon
\end{equation}

\noindent where....

As is the case of cross-sectional data, it is possible identify spatial autocorrelation in the residuals of Model \ref{eq:sur-sim}, and several Lagrange Multipliers tests have been developed to test the null hypothesis of spatial independence [see @Lopez2014spatial]. When the null hypothesis is rejected, there are alternative several specifications that have been propose to include spatial effects [@Anselin1988spatial; see also @Anselin2016estimation]. In this paper we consider a spatial SUR model that incorporates a spatial lag of the dependent variable as an explanatory factor. The stack expresion for the spatial lag SUR model (SUR-SLM),

\begin{equation}
\label{eq:sur-slm}
\bf{A}y = X \beta + \epsilon \\
\epsilon =N(0,\Omega)
\end{equation}

\noindent where A =$I_{TN}-\bf{\Lambda} \otimes W$ with $\bf{\Lambda} = diag(\lambda_1, \cdots, \lambda_T)$.

This specification assumes that incidence in a province ($y_st$) at time $t$ is partially determined by the weighted average ($Wy_{st}$) of incidence in neighbouring provinces. Parameters of spatial dependence are estimated for each time period $\lambda_t$ and identifies the intensity and the sign of the impacts of neighbourhood. It is possible test the null hypotheis of identtical levels of spatial dependence ($\lambda_i=\lambda_j, \forall i,j$). The correspond Wald test is available in the R package **spsur**

The SUR-SLM model can be estimated using maximum likelihood (@Lopez2014spatial) or instrumental variables (@Minguez2019).

Analysis and Results
===================

## Exploratory Data Analysis

We begin with the exploratory analysis of the data.

Figure \ref{fig:weekly-average-incidence-map} shows the geographical variation in the incidence of the disease, as well as its temporal progression in weekly averages. Our analysis begins on March 13, which is the first date when every province had reported at least one case of COVID-19. It can be seen that the highest incidence at this early date was in the provice of Álava, in the North of Spain. While not the most populous province, with a population of only 331,549, Álava has the highest GDP per capita of all provinces. Vitoria, its main city, is the capital of the Basque country and has been the focus of efforts to develop a "Global Basque City" [@Meijers2008strategic], along with San Sebastian and Bilbao. The other early focus of the pandemic in Spain was Madrid, which is the most populous province in the country and has the second highest GDP per capita after Álava. The early outbreaks in these two provinces can be traced throughout the progression of the pandemic over time, although by the end of the period under consideration, other provinces had matched and even surpased their incidence rates, including Segovia and Soria to the north of Madrid, and Ciudad Real and Albacete to the south.

```{r weekly-average-incidence-map, echo=FALSE, fig.height= 9, fig.cap="\\label{fig:weekly-average-incidence-map}Mean weekly incidence of COVID-19 by province, in reported cases by 100,000 people (map omits Islas Canarias)", warning=FALSE}
# Retrieve the coordinates of Madrid
madrid_xy <-  st_centroid(provinces_spain %>% filter(province == "Madrid")) %>% st_coordinates()

week11.plot <- covid19_spain %>%
  filter(CCAA != "Canarias") %>%
  group_by(province, week = isoweek(Date)) %>% 
  summarise(ID_INE = first(ID_INE), mean_weekly_incidence = mean(Incidence)) %>%
  filter(week == 11) %>%
  left_join(provinces_spain,
            by = c("ID_INE")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = mean_weekly_incidence)) +
  scale_fill_distiller(name = "Mean Weekly Incidence", 
                       palette = "Reds", 
                       direction = 1) +
  ggtitle("March 13 - March 15") +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  geom_sf_text(label = "Madrid", 
               x = madrid_xy[1] - 4.9, 
               y = madrid_xy[2] + 0.6,
               family = "serif") +
  geom_segment(x = madrid_xy[1] - 3.4, 
               y = madrid_xy[2] + 0.5, 
               xend = madrid_xy[1], 
               yend = madrid_xy[2])

week12.plot <- covid19_spain %>%
  filter(CCAA != "Canarias") %>%
  group_by(province, week = isoweek(Date)) %>% 
  summarise(ID_INE = first(ID_INE), mean_weekly_incidence = mean(Incidence)) %>%
  filter(week == 12) %>%
  left_join(provinces_spain,
            by = c("ID_INE")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = mean_weekly_incidence)) +
  scale_fill_distiller(name = "Mean Weekly Incidence", 
                       palette = "Reds",
                       direction = 1) +
  ggtitle("March 16 - March 22") +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  geom_sf_text(label = "Madrid", x = madrid_xy[1] - 4.9, y = madrid_xy[2] + 0.6) +
  geom_segment(x = madrid_xy[1] - 3.4, y = madrid_xy[2] + 0.5, xend = madrid_xy[1], yend = madrid_xy[2])

week13.plot <- covid19_spain %>%
  filter(CCAA != "Canarias") %>%
  group_by(province, week = isoweek(Date)) %>% 
  summarise(ID_INE = first(ID_INE), mean_weekly_incidence = mean(Incidence)) %>%
  filter(week == 13) %>%
  left_join(provinces_spain,
            by = c("ID_INE")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = mean_weekly_incidence)) +
  scale_fill_distiller(name = "Mean Weekly Incidence", 
                       palette = "Reds", 
                       direction = 1) +
  ggtitle("March 23 - March 29") +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  geom_sf_text(label = "Madrid",
               x = madrid_xy[1] - 4.9, 
               y = madrid_xy[2] + 0.6,
               family = "serif") +
  geom_segment(x = madrid_xy[1] - 3.4, 
               y = madrid_xy[2] + 0.5, 
               xend = madrid_xy[1], 
               yend = madrid_xy[2])

week14.plot <- covid19_spain %>%
  filter(CCAA != "Canarias") %>%
  group_by(province, week = isoweek(Date)) %>% 
  summarise(ID_INE = first(ID_INE), mean_weekly_incidence = mean(Incidence)) %>%
  filter(week == 14) %>%
  left_join(provinces_spain,
            by = c("ID_INE")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = mean_weekly_incidence)) +
  scale_fill_distiller(name = "Mean Weekly Incidence", 
                       palette = "Reds", 
                       direction = 1) +
  ggtitle("March 30 - April 5") +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  geom_sf_text(label = "Madrid", 
               x = madrid_xy[1] - 4.9, 
               y = madrid_xy[2] + 0.6,
               family = "serif") +
  geom_segment(x = madrid_xy[1] - 3.4, 
               y = madrid_xy[2] + 0.5, 
               xend = madrid_xy[1], 
               yend = madrid_xy[2])

week15.plot <- covid19_spain %>%
  filter(CCAA != "Canarias") %>%
  group_by(province, week = isoweek(Date)) %>% 
  summarise(ID_INE = first(ID_INE), mean_weekly_incidence = mean(Incidence)) %>%
  filter(week == 15) %>%
  left_join(provinces_spain,
            by = c("ID_INE")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = mean_weekly_incidence)) +
  scale_fill_distiller(name = "Mean Weekly Incidence", 
                       palette = "Reds", 
                       direction = 1) +
  ggtitle("April 6 - April 11") +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  geom_sf_text(label = "Madrid", 
               x = madrid_xy[1] - 4.9, 
               y = madrid_xy[2] + 0.6,
               family = "serif") +
  geom_segment(x = madrid_xy[1] - 3.4, 
               y = madrid_xy[2] + 0.5, 
               xend = madrid_xy[1], 
               yend = madrid_xy[2])

grid.arrange(week11.plot, week12.plot, week13.plot, week14.plot, week15.plot, nrow = 3)
```

To visualize the distribution of temperature and humidity we aggregate the provinces by Autonomous Community. In Figure \ref{fig:descriptives-temperature} the communities have been sorted by latitude, so that Principado de Asturias is the northernmost community, and Canarias the southernmost. There is a relatively wide range of values both within and between provinces over the 30 day period examined. The top panel of the figure shows the distribution of mean temperature. The lowest mean temperature for a community on any given day is 2.8C, and the highest is 22.4C for a range of approximately 20 degrees. Likewise, there is a great deal of variability in humidity, as seen in the bottom panel of the figure, where the lowest mean humidity for any community is 48.3 and the highest is 99.6. The actual values for the provinces display somewhat more variability even.

```{r descriptives-temperature, echo = FALSE, fig.height=7, fig.cap="\\label{fig:descriptives-temperature} Distribution of mean temperatures and humidities in the Autonomous Communities in Spain between March 12, 2020 and April 11, 2020. The Autonomous Communities have been sorted by latitude, with communities to the left being the northermost, and to the right the southernmost"}
# Summary of temperature by Autonomous Community
temp.summary <- covid19_spain %>% 
  group_by(CCAA, Date) %>%
  summarize(Value = mean(Mean_Temp)) %>%
  mutate(Variable = "Mean Temperature in Community")

# Summary of humidity by Autonomous Community
humidity.summary <- covid19_spain %>% 
  group_by(CCAA, Date) %>%
  summarize(Value = mean(Humidity)) %>%
  mutate(Variable = "Mean Humidity in Community")

# Bind tables
env.summary <- rbind(temp.summary, humidity.summary) %>%
  mutate(Variable = factor(Variable, 
                           levels = c("Mean Temperature in Community",
                                      "Mean Humidity in Community")))

# Plot
ggplot(data = env.summary, 
       aes(x = CCAA,
           y = Value)) +
  geom_boxplot() +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("Autonomous Community (sorted from north to south)") +
  ylab("Value") +
  facet_wrap(~ Variable, nrow = 2, 
             scales = "free_y")
```

```{r control-map, echo=FALSE, fig.height= 9, fig.cap="\\label{fig:control-map}Spatial distribution of control variables by province (maps omits Islas Canarias)"}
gdppc.plot <- provinces_spain %>%
  filter(CCAA != "Canarias") %>%
  ggplot() +
  geom_sf(aes(fill = GDPpc)) +
  scale_fill_distiller(name = "GDP per capita", 
                       palette = "Blues", 
                       direction = 1) +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom")

# Older adults:
older.plot <- provinces_spain %>% 
  filter(CCAA != "Canarias") %>%
  ggplot() +
  geom_sf(aes(fill = as.numeric(Older))) +
  scale_fill_distiller(name = "Percentage of older adults", 
                       palette = "Blues", 
                       direction = 1) +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom")

# Population density:
density.plot <- provinces_spain %>% 
  filter(CCAA != "Canarias") %>%
  ggplot() +
  geom_sf(aes(fill = as.numeric(Density))) +
  scale_fill_distiller(name = "Population Density", 
                       palette = "Blues", 
                       direction = 1) +
  theme_tufte() +
  theme(axis.text = element_blank(),
        legend.position = "bottom")

grid.arrange(gdppc.plot, older.plot, density.plot, nrow = 3)
```

Figure \ref{fig:daily-correlations} shows the distribution of daily correlations of the independent variables with incidence of COVID-19, after log-transforming all variables. It can be seen there that the correlation of GDPpc and temperature (in its three definitions) have the strongest positive and negative correlations with incidence, respectively. Percentage of older adults displays somewhat weaker negative correlations with incidence, as does density. It can be seen that the humidity variable, in its three forms, tends to be possitively correlated with incidence of COVID-19.

```{r daily-correlations, echo=FALSE, fig.cap="\\label{fig:daily-correlations}Distribution of daily correlations of the independent variables with daily incidence of COVID-19 (all variables have been log-transformed)"}
covid19_spain %>% 
  group_by(Date) %>%
  summarize(Older = cor(log(Older), log(Incidence)),
            GDPpc = cor(log(GDPpc), log(Incidence)),
            Density = cor(log(Density), log(Incidence)),
            Humidity_lag8 = cor(Humidity_lag8, Incidence),
                        Humidity_lag11 = cor(log(Humidity_lag11), log(Incidence)),
                        Humidity_lag11w = cor(log(Humidity_lag11w), log(Incidence)),
            Mean_Temp_lag8 = cor(log(Mean_Temp_lag8), log(Incidence)),
            Mean_Temp_lag11 = cor(log(Mean_Temp_lag11), log(Incidence)),
            Mean_Temp_lag11w = cor(log(Mean_Temp_lag11w), log(Incidence))) %>%
  select(-Date) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Correlation") %>%
  mutate(Variable = factor(Variable,
                           levels = c("GDPpc", 
                                      "Older", 
                                      "Density", 
                                      "Humidity_lag8",
                                      "Humidity_lag11",
                                      "Humidity_lag11w",
                                      "Mean_Temp_lag8",
                                      "Mean_Temp_lag11",
                                      "Mean_Temp_lag11w"))) %>%
  ggplot(aes(x = Variable, y = Correlation)) +
  geom_boxplot() +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

### SUR Models

```{r data-preparation-for-modelling, include=FALSE}
# Organize data for SUR modelling 
GPanel <- plm::pdata.frame(covid19_spain %>%
                             select(province, 
                                    Date,
                                    Incidence, 
                                    Median_Age,
                                    Male2Female,
                                    Older,
                                    GDPpc,
                                    Density,
                                    Transit,
                                    Mean_Temp_lag8,
                                    Humidity_lag8,
                                    Sunshine_Hours_lag8,
                                    Mean_Temp_lag11,
                                    Humidity_lag11,
                                    Sunshine_Hours_lag11,
                                    Mean_Temp_lag11w,
                                    Humidity_lag11w,
                                    Sunshine_Hours_lag11w),
                           c("province","Date"))

```

```{r spatial-weights, include=FALSE}
# Create spatial weights matrix: 
Wmat <- provinces_spain %>%
  #drop_na() %>%
  as("Spatial") %>%
  poly2nb(queen = FALSE) %>%
  nb2mat(zero.policy = TRUE)

Wmat <- (Wmat > 0) * 1

# Join the two provinces in Canarias
Wmat[which(provinces_spain$province == "Palmas(Las)"), 
     which(provinces_spain$province == "Santa Cruz de Tenerife")] <- 1
Wmat[which(provinces_spain$province == "Santa Cruz de Tenerife"), 
     which(provinces_spain$province == "Palmas(Las)")] <- 1

# 'Paises Catalans'
#n = 8
Wmat[which(provinces_spain$province == "Barcelona"), 
     which(provinces_spain$province == "Baleares")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Barcelona")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Castellon/Castello")] <- 1 
Wmat[which(provinces_spain$province == "Castellon/Castello"), 
     which(provinces_spain$province == "Baleares")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Tarragona")] <- 1 
Wmat[which(provinces_spain$province == "Tarragona"), 
     which(provinces_spain$province == "Baleares")] <- 1
miW <- Wmat/rowSums(Wmat)

# Convert to listw
listw <- mat2listw(Wmat,style = "W")
```

```{r formulas, include=FALSE}
# Define formulas with three different lagged variables:
formula_lag8 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag8) +
  log(Mean_Temp_lag8)

formula_lag11 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11) +
  log(Mean_Temp_lag11)

formula_lag11w <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11w) +
  log(Mean_Temp_lag11w)
```

```{r model-restrictions, include=FALSE}
T <- max(covid19_spain$Date) - min(covid19_spain$Date) + 1 # Recall that T is the number of days, i.e., time periods, i.e., equations; add 1 to include the starting day
k <- 7 # Number of independent variables, including the constant
coef_rest <- 2 # Number of restrictions

# nrow is number of equations (time periods) minus 1, times the number of restrictions
# ncol is number of variables times number of equations
R2 <- matrix(0, nrow = (T - 1) * coef_rest, ncol = k * T)

for (i in 1:(T-1)){
  R2[i, 2] <- 1
  R2[i, (2 + i * k)] <- -1
  R2[(i + T - 1), 3] <- 1
  R2[(i + T - 1), (3 + i * k)] <- -1
  # Use if more restrictions are needed
  #R2[(i + T - 1) * 2, 4] <- 1
  #R2[(i + T - 1) * 2, (4 + i * k)] <- -1
}
b2 <- matrix(0, ncol = 21*coef_rest)
```

```{r model1-8-day-moving-average, include=FALSE}
# Model with a lagged 8-day moving average of climatic variables:
sur.slm_lag8 <- spsur::spsurtime(formula = formula_lag8, 
                                 data=GPanel, 
                                 time = GPanel$Date, 
                                 type = "slm", 
                                 fit_method = "3sls", 
                                 listw=  listw,
                                 R = R2,
                                 b = b2)
#summary(sur.slm_lag8)
#print(paste("Pooled R^2 = ", sur.slm_lag8$R2[1]))
```

```{r model2-11-day-moving-average, include=FALSE}
# Model with 11-day moving average of climatic variables:
sur.slm_lag11 <- spsur::spsurtime(formula = formula_lag11, 
                                  data=GPanel, 
                                  time = GPanel$Date, 
                                  type = "slm", 
                                  fit_method = "3sls", 
                                  listw=  listw,
                                 R = R2,
                                 b = b2)
#summary(sur.slm_lag11)
#print(paste("Pooled R^2 = ", sur.slm_lag11$R2[1]))
```

```{r model3-11-day-weighted-moving-average, include=FALSE}
# Model with 11-day weighted moving average of climatic variables:
sur.slm_lag11w <- spsur::spsurtime(formula = formula_lag11w, 
                                   data=GPanel, 
                                   time = GPanel$Date, 
                                   type = "slm", 
                                   fit_method = "3sls", 
                                   listw=  listw,
                                 R = R2,
                                 b = b2)
#summary(sur.slm_lag11w)
#print(paste("Pooled R^2 = ", sur.slm_lag11w$R2[1]))
```

The goodness of fit of the three systems of equations is shown in Figure \ref{fig:goodness-of-fit}.

```{r goodness-of-fit, echo=FALSE, fig.cap="\\label{fig:goodness-of-fit} Goodness of fit of the SUR systems: by date and pooled"}
data.frame(Model1 = sur.slm_lag8$R2, 
           Model2 = sur.slm_lag11$R2,
           Model3 = sur.slm_lag11w$R2) %>%
  slice(2:n()) %>%
  rownames_to_column(var = "Equation") %>%
  mutate(Date = seq(ymd("2020-03-13"), 
                    ymd("2020-04-11"), 
                    by = "days")) %>%
  pivot_longer(cols = starts_with("M"), names_to = "Model", values_to = "R2") %>%
  mutate(Model = factor(Model, 
                        levels = c("Model1", "Model2", "Model3"),
                        labels = c("Model 1", "Model 2", "Model 3"))) %>%
  ggplot(aes(x = Date, y = R2, color = Model, shape = Model)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Model 1" = "blue", "Model 2" = "orange", "Model 3" = "green") ) + 
  annotate(geom = "text", 
           label = c(paste("Model 1: pooled-R2=", 
                          round(sur.slm_lag8$R2[1], 4)),
                     paste("Model 2: pooled-R2=", 
                          round(sur.slm_lag11$R2[1], 4)),
                     paste("Model 3: pooled-R2=", 
                          round(sur.slm_lag11w$R2[1], 4))),
            x = as_date("2020-04-01"),
            y = c(0.3, 0.27, 0.24),
           family = "serif",
           size = 3.5) +
  theme_tufte()

```

Summary of best model.
```{r summary-best-model, echo=FALSE}
data.frame(Coefficient = sur.slm_lag11$coefficients, 
                   pval = dnorm(abs(sur.slm_lag11$coefficients/sur.slm_lag11$rest.se))) %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = c("Intercept", 
                      "log(GDPpc)", 
                      "log(Older)",
                      "log(Density)",
                      "Transit",
                      "log(Humidity)",
                      "log(Temperature)",
                      rep(c("Intercept",
                            "log(Density)", 
                            "Transit", 
                            "log(Humidity)", 
                            "log(Temperature)"), 29)),
         Variable = factor(Variable, 
                           levels = c("Intercept", 
                                      "log(GDPpc)", 
                                      "log(Older)", 
                                      "log(Density)",
                                      "Transit", 
                                      "log(Humidity)",
                                      "log(Temperature)"))) %>%
  group_by(Variable) %>%
  summarize(Min = min(Coefficient), 
            Mean = mean(Coefficient), 
            Max = max(Coefficient),
            pns = sum(pval > 0.10), # Number of coefficients with p > 0.1
            p10 = sum(pval <= 0.10) - sum(pval <= 0.05), # Number of coefficients with  0.1 <= pval < 0.05
            p05 = sum(pval <= 0.05)) %>% # Number of coefficients with pval <=5 
  mutate(Note = c("Non-constrained", 
                  "Constrained", 
                  "Constrained", 
                  "Non-constrained", 
                  "Non-constrained",
                  "Non-constrained",
                  "Non-constrained")) %>%
  kable("html",
        booktabs = TRUE,
        digits = 3,
        col.names = c("Variable", "Min", "Mean", "Max", "p > 0.10", "0.10 <= p < 0.05", "p <= 0.05", "Note"),
        caption = "\\label{tab:summary-best-model}Summary of estimation results best model (lagged 11-day moving average)") %>%
  kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 1, "Estimates" = 3, "Significance" = 3, " " = 1)) %>%
  footnote(general = c("Significance: This is the number of coefficients with p-values as indicated",
                       "Non-constrained: coefficient was allowed to vary across equations",
                       "Constrained: coefficient as constant across equations"))
```

## Discussion

Figure \ref{fig:delta-time} shows the temporal evolution of the spatial autocorrelation coefficient ($\rho$).

```{r delta-time, echo=FALSE, fig.cap="\\label{fig:delta-time}Temporal evolution of spatial autocorrelation coefficient"}
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Deltas = sur.slm_lag11$deltas,
           tvalue = sur.slm_lag11$deltas/sur.slm_lag11$deltas.se) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Deltas, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

Figure \ref{fig:beta-intercept-time} shows the temporal evolution of the intercept.

```{r beta-intercept-time, echo=FALSE, fig.cap="\\label{fig:beta-intercept-time}Temporal evolution of intercept"}
#Intercept
n = 1
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

Figure \ref{fig:beta-density-time} shows the temporal evolution of the coefficient for $\log(Density)$.

```{r beta-density-time, echo=FALSE, , fig.cap="\\label{fig:beta-density-time}Temporal evolution of coefficient for log(Density)"}
#Density:
n = 2 # Although this is the fourth coefficient, after intercept, log(GDPpc), and log(Older), we start at 2 here because the two constrained coefficients are removed from the matrix of coefficients 
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

Figure \ref{fig:beta-transit-time} shows the temporal evolution of the coefficient for $Transit$.

```{r beta-transit-time, echo=FALSE, fig.cap="\\label{fig:beta-transit-time}Temporal evolution of coefficient for Transit"}
n = 3
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

Figure \ref{fig:beta-humidity-time} shows the temporal evolution of the coefficient for $\log(Humidity)$.

```{r beta-humidity-time, echo=FALSE, fig.cap="\\label{fig:beta-humidity-time}Temporal evolution of coefficient for log(Humidity)"}
n = 4
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

Figure \ref{fig:beta-density-time} shows the temporal evolution of the coefficient for $\log(Temperature)$.

```{r beta-temperature-time, echo=FALSE, fig.cap="\\label{fig:beta-temperature-time}Temporal evolution of coefficient for log(Temperature)"}
n = 5
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

## Concluding Remarks

More words go here.

## Acknowledgments

The following `R` packages were used in the course of this investigation and the authors wish to acknowledge their developers: `ggthemes` [@Arnold2019], `gridExtra` [@Auguie2017gridextra], `kableExtra` [@Zhu2019], `knitr` [@Xie2014; @Xie2015], `lubridate` [@Grolemund2011dates], `plm` [@Millo2017robust], `rticles` [@Allaire2020], `sf` [@Pebesma2018], `spdep` [@Bivand2013], spsur [@Angulo2020spsur] `tidyverse` [@Wickham2019], `units` [@Pebesma2016].

## References