### 2020-05-06 Visualization-Contagion/Visualizing-Contagion.Rmd

---
title: "Visualize contagion"
output:
  html_document:
    df_print: paged
---

In this notebook I'd like to visualize the pattern of inter-provincial contagion. My idea is to get the fitted values of the model minus the effect of the covariates, to isolate the lagged dependent variable, and then do an animation with that.

## Preliminaries

Load packages:
```{r load-packages, message=FALSE}
# Download and install these two packages:
# covid19env
# spsur
library(covid19env)
library(gganimate)
#library(ggthemes)
#library(gridExtra)
#library(kableExtra)
library(lubridate)
library(plm)
#library(plotly)
library(sf)
library(spdep)
library(spsur)
library(tidyverse)
library(units)
```

Load datasets:
```{r load-data, cache=TRUE}
# Load data from package `covid19env`
data("covid19_spain")
data("provinces_spain")
```

Preprocess the data:
```{r data-preparation, cache=TRUE}
# Convert GDP per capita to thousands of euros:
provinces_spain <- provinces_spain %>%
  mutate(GDPpc = GDPpc/1000)

# Join provincial data to incidence data and convert to simple features:
covid19_spain <- covid19_spain %>% 
  left_join(provinces_spain %>% st_drop_geometry(),
            by = c("province", "CCAA", "ID_INE"))
```

## The contagion effect

Consider an estimated model of the following form:
$$
(I -  \hat{\rho}W)\hat{Y} = X\hat{\beta}
$$

Expand:
$$
\hat{Y} - \hat{\rho}W\hat{Y} = X\hat{\beta}
$$

Rearrange:
$$
\hat{\rho}W\hat{Y} = \hat{Y} - X\hat{\beta}
$$

\noindent where $\hat{Y}$ is the prediction of incidence in the provinces, and $X\hat{\beta}$ is the estimated impact on incidence due to all $X$ factors. The right-hand side of the equation is the estimated incidence discounted by its own factors.

The term on the right hand side is the contribution to $\hat{Y}$ due to incidence in the neighboring provinces. In other words, this is the contagion effect.

When the estimated incidence $\hat{Y}$ is greater than $X\hat{\beta}$, this means that the province has a greater incidence of the disease than what its own conditions explain. In this case $\hat{\rho}W\hat{Y}$ is positive, and contagion contributes to increase the incidence (the province tends to be on the receiving end of contagion).

On the other hand, when the estimated incidence $\hat{Y}$ is less than $X\hat{\beta}$, the province has lower incidence than what its own conditions explain. Now, $\hat{\rho}W\hat{Y}$ is negative, which means that the incidence is reduced by lower rates of incidence in their neighborhood.

Finally, when the estimated incidence $\hat{Y}$ is equal to $X\hat{\beta}$, the incidence in the province is explained completely by its own conditions.

By mapping this term, we can explore how vulnerable to contagion, or contagious, the provinces were over time. Remember, positive means on the receiving end (or vulnerable) of contagion, red means contagious.

## Model

Organize data for SUR modelling: 
```{r data-preparation-for-modelling, cache = TRUE}
GPanel <- plm::pdata.frame(covid19_spain %>%
                             select(province, 
                                    Date,
                                    Incidence, 
                                    Median_Age,
                                    Male2Female,
                                    Older,
                                    GDPpc,
                                    Density,
                                    Transit,
                                    Mean_Temp_lag8,
                                    Humidity_lag8,
                                    Sunshine_Hours_lag8,
                                    Mean_Temp_lag11,
                                    Humidity_lag11,
                                    Sunshine_Hours_lag11,
                                    Mean_Temp_lag11w,
                                    Humidity_lag11w,
                                    Sunshine_Hours_lag11w),
                           c("province","Date"))

```

Create spatial weights matrix: 
```{r spatial-weights, cache = TRUE}

Wmat <- provinces_spain %>%
  #drop_na() %>%
  as("Spatial") %>%
  poly2nb(queen = FALSE) %>%
  nb2mat(zero.policy = TRUE)

Wmat <- (Wmat > 0) * 1

# Join the two provinces in Canarias
Wmat[which(provinces_spain$province == "Palmas(Las)"), 
     which(provinces_spain$province == "Santa Cruz de Tenerife")] <- 1
Wmat[which(provinces_spain$province == "Santa Cruz de Tenerife"), 
     which(provinces_spain$province == "Palmas(Las)")] <- 1

# 'Paises Catalans'
#n = 8
Wmat[which(provinces_spain$province == "Barcelona"), 
     which(provinces_spain$province == "Baleares")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Barcelona")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Castellon/Castello")] <- 1 
Wmat[which(provinces_spain$province == "Castellon/Castello"), 
     which(provinces_spain$province == "Baleares")] <- 1
Wmat[which(provinces_spain$province == "Baleares"), 
     which(provinces_spain$province == "Tarragona")] <- 1 
Wmat[which(provinces_spain$province == "Tarragona"), 
     which(provinces_spain$province == "Baleares")] <- 1
miW <- Wmat/rowSums(Wmat)

# Convert to listw
listw <- mat2listw(Wmat,style = "W")
```

Define formula for model:
```{r formulas, cache = TRUE}
formula_lag11 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11) +
  log(Mean_Temp_lag11) +
  log(Sunshine_Hours_lag11 + 0.1)
```

Add restrictions to model for temporally fixed coefficients:
```{r model-restrictions, cache = TRUE}
T <- max(covid19_spain$Date) - min(covid19_spain$Date) + 1 # Recall that T is the number of days, i.e., time periods, i.e., equations; add 1 to include the starting day
k <- 8 # Number of independent variables, including the constant
coef_rest <- 2 # Number of restrictions

# nrow is number of equations (time periods) minus 1, times the number of restrictions
# ncol is number of variables times number of equations
R2 <- matrix(0, nrow = (T - 1) * coef_rest, ncol = k * T)

for (i in 1:(T-1)){
  R2[i, 2] <- 1
  R2[i, (2 + i * k)] <- -1
  R2[(i + T - 1), 3] <- 1
  R2[(i + T - 1), (3 + i * k)] <- -1
  # Use if more restrictions are needed
  #R2[(i + T - 1) * 2, 4] <- 1
  #R2[(i + T - 1) * 2, (4 + i * k)] <- -1
}
b2 <- matrix(0, ncol = (T - 1) * coef_rest)
```

Estimate model:
```{r model2-11-day-moving-average, cache = TRUE}
# Model with 11-day moving average of climatic variables:
sur.slm_lag11 <- spsur::spsurtime(formula = formula_lag11, 
                                  data=GPanel, 
                                  time = GPanel$Date, 
                                  type = "slm", 
                                  fit_method = "3sls", 
                                  listw=  listw,
                                  R = R2,
                                  b = b2)
#summary(sur.slm_lag11)
#print(paste("Pooled R^2 = ", sur.slm_lag11$R2[1]))
```

## Prepare results for analysis

To analyze the contagion effect we need to calculate it. This can be done in two different ways. Directly, by computing this term:

$$
\hat{\rho}W\hat{Y}
$$

The estimated autocorrelation coefficient $\hat{\rho}$ (by date) and the fitted values $\hat{Y}$ are part of the output of the estimation procedure.

As an alternative, the contagion effect can be calculated indirectly, as:
$$
\hat{Y} - X\hat{\beta}
$$

The fitted values $\hat{Y}$ and the coefficients $\hat{\beta}$ (by date) are part of the output of the estimation procedure. Here, I take the second approach, which avoid having to work with matrix $W$ in the calculations.

Extract and organize coefficients:
```{r}
Coefficients <- data.frame(Coefficients = sur.slm_lag11$coefficients,
                           Variable = c(c("Intercept",
                                        "b_GDPpc",
                                        "b_Older",
                                        "b_Density",
                                        "b_Transit",
                                        "b_Humidity",
                                        "b_Temperature",
                                        "b_Sunshine"),
                                        rep(c("Intercept",
                                              "b_Density",
                                              "b_Transit",
                                              "b_Humidity",
                                              "b_Temperature",
                                              "b_Sunshine"),
                                            29)),
                           Date = c(rep(ymd("2020-03-13"), 8),
                                    rep(seq(ymd("2020-03-14"),
                                        ymd("2020-04-11"),
                                        by = "days"),
                                        each = 6)))  %>%
  pivot_wider(names_from = "Variable", 
              values_from = "Coefficients")  %>%
  mutate(b_GDPpc = max(sur.slm_lag11$coefficients[2]),
         b_Older = max(sur.slm_lag11$coefficients[3]))
```

Join coefficients with data table:
```{r}
covid19_spain <- covid19_spain %>% left_join(Coefficients, by = "Date")
```

Calculate $\hat{Y}$ minus $X\hat{\beta}$:
```{r}
Contagion <- covid19_spain %>%
  mutate(fitted.values = sur.slm_lag11$fitted.values,
         X_beta = Intercept + 
           b_GDPpc * log(GDPpc) + 
           b_Older * log(Older) + 
           b_Density * log(as.numeric(Density)) +
           b_Transit * Transit +
           b_Humidity * log(Humidity_lag11) +
           b_Temperature * log(Mean_Temp_lag11) + 
           b_Sunshine * log(Sunshine_Hours_lag11 + 0.01), 
         Contagion = fitted.values - X_beta)
```

Select variables from `Contagion` dataframe:
```{r}
Contagion <- Contagion %>%
  select(ID_INE, province, CCAA, Contagion, Date) %>%
  mutate(Day = factor(rep(1:50, 30)))
```

Next, join geometry and convert to sf:
```{r}
Contagion <- Contagion %>%
  left_join(provinces_spain %>% 
              select(ID_INE, geometry),
            by = "ID_INE") %>%
  st_as_sf()
```

Create a choropleth plot of the provinces with the contagion effect ($\hat{\rho}W\hat{Y}$):
```{r}
p1 <- ggplot() + 
  geom_sf(data = Contagion %>% 
            filter(CCAA != "Canarias"), 
          aes(fill = Contagion, 
              group = interaction(factor(ID_INE), 
                                  Date))) +
  scale_fill_gradient2() + 
  theme_bw()
```

Create an animation using `gganimate`:
```{r}
anim1 <-  p1 +
  labs(title = "Date: {frame_time}") +
  transition_time(Date)
```

Render animation:
```{r cache=TRUE, eval=FALSE}
animate(anim1, fps = 2, end_pause = 5)
```

What general trends to we see?

Now plot the trajectory of the contagion term by province:
```{r}
ggplot(data = Contagion, aes(x = Date, y = Contagion, group = province)) +
  geom_line(color = "gray") +
  geom_line(data = Contagion %>% 
              filter(province == "Araba/alava"), 
            color = "orange",
            size = 1) +
  geom_line(data = Contagion %>% 
              filter(province == "Madrid"), 
            color = "purple",
            size = 1) +
  labs(title = "Contagion effect by province by date",
              #subtitle = "Plot by date",
              caption = "Alava is the orange line; Madrid is the purple line") +
  theme_bw()
```

Where could this analysis be taken? LISA analysis by date? Spatial Markov Chains? Cluster membership?

## LISA analysis of contagion

Here I calculate LISA statistics for `Contagion` on each date:
```{r}
junk <- Contagion %>%
  group_by(Date) %>%
  summarize(Ii =, 
            P_Ii = )
localmoran(Contagion$Contagion[Contagion$Date == "2020-03-13"], listw)[,5]
```

LISA statistic calculated for each cross-section
```{r}
Ii <- Contagion %>% 
  group_by(Date) %>%
  group_modify(~ {(localmoran(.x$Contagion, listw)[,1]) %>%
      enframe()}) %>%
  ungroup()  %>%
  transmute(Date,
            ID_INE = rep(c(1:50), 30),
            Ii = value)
```

LISA's p-value:
```{r}
P_Ii <- Contagion %>% 
  group_by(Date) %>%
  group_modify(~ {(localmoran(.x$Contagion, listw)[,5]) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(Date,
            ID_INE = rep(c(1:50), 30),
            P_Ii = value)
```

Join the results of the LISA analysis to the dataframe:
```{r}
Contagion <- Contagion %>%
  left_join(Ii, by = c("Date" = "Date", "ID_INE" = "ID_INE")) %>%
  left_join(P_Ii, by = c("Date" = "Date", "ID_INE" = "ID_INE"))
```

Label provinces by whether they are positive or negative, and whether they are significant or not:
```{r}
Contagion <- Contagion %>%
  mutate(LISA = case_when(Ii <= 0 & P_Ii <= 0.05 ~ "Hot Spot",
                          Ii <= 0 & P_Ii > 0.05 ~ "NS (Hot)",
                          Ii > 0 & P_Ii <= 0.05 ~ "Cold Spot",
                          Ii > 0 & P_Ii > 0.05 ~ "NS (Cold)"),
         LISA = factor(LISA, 
                       levels = c("Hot Spot",
                                  "NS (Hot)",
                                  "NS (Cold)",
                                  "Cold Spot"),
                       ordered = TRUE))
```

Create a choropleth plot of the provinces based on the results of the LISA analysis:
```{r}
p2 <- ggplot() + 
  geom_sf(data = Contagion %>% 
            filter(CCAA != "Canarias"), 
          aes(fill = LISA, 
              group = interaction(factor(ID_INE), 
                                  Date))) +
  scale_fill_manual(values = c("Hot Spot" = "red", 
                               "NS (Hot)" = "pink", 
                               "NS (Cold)" = "royalblue", 
                               "Cold Spot" = "royalblue4")) + 
  theme_bw()
```

Create an animation using `gganimate`:
```{r}
anim2 <-  p2 +
  labs(title = "Date: {frame_time}") +
  transition_time(Date)
```

Render animation:
```{r cache=TRUE, eval=FALSE}
animate(anim2, fps = 2, end_pause = 5)
```

## Spatial Markov Chains

I used to have Matlab code for spatial markov chains (SMC). It is gone. Here' I will try to reproduce it but now in `R`.

The first step to implement SMC is to discretize the variable of interest. This can be done in two ways: within period quantiles, and pooled quantiles for the whole time series. According to [Rey and Kang](https://pysal.org/notebooks/explore/giddy/Markov_Based_Methods.html), these two forms of doing the analysis give different interpretations: pooled quantiles are absolute trends, whereas within period quantiles are relative dynamics.

Here, we label each case by their pooled quintile (`q_p`):
```{r}
Contagion <- Contagion %>%
  mutate(q_p = cut(Contagion, 
                   quantile(Contagion, c(0, 0.2, 0.4, 0.6, 0.8, 1)), 
                   labels = c("High Contagion", 
                              "Medium-High Contagion", 
                              "Medium Contagion", 
                              "Medium-Low Contagion", 
                              "Low Contagion"))) %>%
  replace_na(replace = list(q_p = "High Contagion"))
```

To label each case by their within day quintile (`q_d`) we need to summarize the quintiles by date:
```{r}
quintiles_by_date <- Contagion %>%
  st_drop_geometry() %>%
  group_by(Date) %>%
  summarize(x20 = quantile(Contagion, 0.2),
            x40 = quantile(Contagion, 0.4),
            x60 = quantile(Contagion, 0.6),
            x80 = quantile(Contagion, 0.8))
```

Join the quintiles by date to the dataframe:
```{r}
Contagion <- Contagion %>%
  left_join(quintiles_by_date, by = "Date")
```

Label the observations according to their values of `Contagion` and the quintiles by date:
```{r}
Contagion <- Contagion %>%
  mutate(q_d = case_when(Contagion <= x20 ~ "High Contagion",
                         Contagion > x20 & Contagion <= x40 ~ "Medium-High Contagion",
                         Contagion > x40 & Contagion <= x60 ~ "Medium Contagion",
                         Contagion > x60 & Contagion <= x80 ~ "Medium-Low Contagion",
                         Contagion > x80 ~ "Low Contagion"),
         q_d = factor(q_d, 
                      levels = c("High Contagion", 
                              "Medium-High Contagion", 
                              "Medium Contagion", 
                              "Medium-Low Contagion", 
                              "Low Contagion"),
                      ordered = TRUE))
```