### 2020-05-06 covid19env/data_processing/01-Preprocess-meteo-data.Rmd

---
title: "Download climatic data from AEMET"
author: "Fernando"
date: "4/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,collapse = TRUE,message = FALSE)
```

```{r libraries}
library(meteoland)
#devtools::install_github("SevillaR/aemet")
library(aemet)
library(lubridate)
library(nngeo)
library(sf)
library(kableExtra)
library(tidyverse)
```

## Load data:

```{r}
load("provinces_spain.RData")
```

## Geometry of Meteo Stations

```{r plot-stations}
ggplot(data = provinces_spain) +
  geom_sf() +
  theme_bw(base_size=6)
```

Load RData file (this is the outcome of the file `00-Download-meteo-data-from-AEMET.Rmd`; check that file for details on how these data were downloaded):
```{r}
load("meteo_data.RData")
```

Convert `meteo_data` to simple features:
```{r}
meteo_data.sf <- meteo_data %>% st_as_sf(coords = c("Lon", "Lat"), crs = 4326)
```

Join the provincial identifier to calculate the meteorological variables as the average of stations in province:
```{r}
meteo_data_joined = st_join(meteo_data.sf, provinces_spain["ID_INE"])
```

Number of stations per day per province:
```{r}
meteo_data_joined %>% 
  group_by(ID_INE, date) %>% 
  summarize(n = n()) %>%
  summary()
```

Provincial averages will use between 1 and 11 stations.

Obtain provincial averages:
```{r}
meteo_data_province <- meteo_data_joined %>% 
  st_drop_geometry() %>%
  group_by(ID_INE, date) %>%
  summarize(Precipitation = mean(Precipitation, na.rm = TRUE),
            MeanTemperature = mean(MeanTemperature, na.rm = TRUE),
            MinTemperature = mean(MinTemperature, na.rm = TRUE),
            MaxTemperature = mean(MaxTemperature, na.rm = TRUE),
            SunshineHours = mean(SunshineHours, na.rm = TRUE))
```

Summarize the dataframe with the provincial daily averages of the meteo variables: 
```{r}
summary(meteo_data_province)
```
There are some NAs, due to some stations not having been read.

Load the dataset that Fernando produced originally:
```{r}
load("COVID-19-Datos_Climaticos_Spain-Extendido.RData")
```

Arrange the hours of sunshine:
```{r}
# Hours of sunshine 
Sun <- Sun %>% select(-ID, -provincia) %>%
  pivot_longer(starts_with("S"), 
               names_to = "Date",
               values_to = "Sunshine_Hours") %>%
  mutate(Date = rep(seq(dmy("01-03-2020"), 
                        dmy("11-04-2020"), 
                        by = "days"), 
                    50))
```

Join to interpolated meteo variables:
```{r}
meteo_data_province <- meteo_data_province %>% 
  left_join(Sun, by = c("ID_INE" = "ID_INE", "date" = "Date"))
```

Replace NAs with the values from Fernando's table:
```{r}
meteo_data_province <- meteo_data_province %>%
  mutate(SunshineHours = ifelse(is.na(SunshineHours), Sunshine_Hours, SunshineHours))
```

Drop meteo data not assigned to a province:
```{r}
meteo_data_province <- meteo_data_province %>%
  drop_na(ID_INE)
```

Summarize again after dropping stations that were not used for any province:
```{r}
summary(meteo_data_province)
```

Tidy up and save:
```{r}
meteo_data_province <- meteo_data_province %>%
  select(-Sunshine_Hours)
```

```{r}
summary(meteo_data_province)
```

What are the NAs in Precipitation?
```{r}
meteo_data_province %>%
  filter(is.na(Precipitation))
```

Replace the NAs by zeros:
```{r}
meteo_data_province <- meteo_data_province %>%
  mutate(Precipitation = ifelse(is.na(Precipitation), 0, Precipitation))
```

What are the NAs in Sunshine? These are dates that we are not using so we can ignore.
```{r}
meteo_data_province %>%
  filter(is.na(SunshineHours))
```

```{r}

save(meteo_data_province, file = "meteo_data_province.RData", compress = "xz")
```